Hierarchical Controller Architecture in Software Defined Networks

### 2013-12-16 paper/paper.tex
references
refs

\subsection{One switch abstraction}
\label{subsec:aggr}
A controller aggregates its underlying network and behaves as one logical switch, with huge number of ports. This logical switch supports control plane protocols such as OpenFlow[ref], and can be controlled by an external entity (referred to as parent controller) to install forwarding rules. One switch abstraction has two parts, (i) compressing the underlying network state to represent a single switch, and (ii) translating instructions (forwarding rules) from the abstract switch to the underlying network. We present these separately below.
\subsubsection{Compression}
The primary switch abstraction comes by compressing the network into one node. Given an underlying network for a controller, the abstract switch exposed by it contains all the ports which are connected to switches outside the network. Such ports are referred to as \emph{external} ports. Conversely, ports which are connected to ports in the same network are not exposed (hidden away from abstraction) and are referred to as \emph{internal} ports. \emph{External} ports form the \emph{ingress} and \emph{egress} ports for the abstract switch. The switch abstraction maintains a list of these abstract ports, which can be (uniquely) mapped to the ports of the underlying network. This basic abstraction is sufficient for a simple control application, but we require more features for a rich and complete experience. A few such abstractions are presented below.
\begin{enumerate}
    \item \emph{Host compression}:
    Hosts in the underlying network are considered as external entities, which are exposed as hosts connected to the abstract switch.Ports which connect to hosts are considered to be\emph{external} ports. However, if the number of hosts in the underlying network is too large, exposing a list of all hosts could result in inefficient compression. Host compression can be performed in two ways, depending on the host IP distribution and the network being considered.
    \begin{itemize}
        \item \emph{IP compression} -
        Similar to the forwarding rules in an internet router, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller. This works well only if the assignment of switches to controllers is uniform and in accordance with prefixes.
        \item \emph{Abstract host list} -
        A controller can expose an abstract host or list of hosts to its parent controller. The parent controller places rules for the abstract hosts in the peer controllers. This approach would be similar to LAN, and requires address translation while entering or leaving the underlying network. This is more flexible, but complicated, and can be used only when IP prefix compression fails.
    \end{itemize}
    \item \emph{Abstract flow table}:
        The basic flow table in a controller is comprise of the rules placed by the parent controller. However, for a full OpenFlow functionality, we need to support counters and queries, as in OpenFlow switches. These responses, again, must be an aggregation of the network. To this end, we must be able to efficiently represent the flows from the underlying network and obtain statistics which can be exposed to the parent. We propose a data structure to maintain, aggregate and update view of the flows in section~\ref{subsubsec:aggr-data}.
\end{enumerate}

@@

[one-big-switch ref] and [pallete ref] propose algorithms to combine and translate end-to-end policies and routing policies to optimally place rules in switches. These algorithms can be used by control applications to optimally place rules, combining the abstract flow table and its local routing policies.

@@

\subsection{Hierarchical arrangement}
We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network. Switches form the leaf nodes in the tree, while controllers form the internal nodes. An internal node acts as a switch for its parent and as a controller for its children. A controller's view does not span the whole network but only the network of its children. A node translates instructions (forwarding rules) from its parent controller to the forwarding rules for its underlying network. Similarly, network updates from the underlying network are translated into network updates arising from one switch abstraction of the given controller and are passed on to parent controller.

The abstraction described above represents a controller at any level of the hierarchy. The only difference between a root controller and an internal controller is the absence of incoming OpenFLow messages in the former, while the latter receives them from its parent. This simplifies the control application which can be agnostic to the hierarchical level it belongs to, as long as it can support incoming OpenFlow messages. [TODO-tree figure].

@@

Hyperflow[Hyperflow-ref] has a logically centralized but physically distributed controller architecture. It localizes decision making in individual controllers by passively synchronizing the global network view. The control application is agnostic to the underlying architecture and takes a global decision, though the effect is local. The events are propagated using a publish/subscribe messaging paradigm.

Onix[Onix ref] presents a platform on top of which a network control plane can be implemented as a distributed system. Control planes written within Onix operate on a global view of the network and use basic state distribution primitives provided by the platform. Thus Onix provides a general distributed state management API for control plane implementations, but leaves the decision making and trade-offs among consistency, durability and scalability to the control applications. This makes the design, deployment, modification and verification of control applications difficult and cost prohibitive.

Kandoo[Kandoo-ref] takes a slightly similar approach to our own and presents a two layer hierarchy of controllers: (i) the bottom layer is a group of controllers with no interconnection and no knowledge of the nework-wide state, and (ii) the top layer is a logically centralized controller that maintains network-wide state. Bottom layer runs a local application for its underlying network, while the top layer subscribes to changes from the bottom layer. Top layer view is subscription dependent and the architecture cannot be used in a generic fasion.

On a similar but different note, B4[B4-ref] uses a two level hierarchy, where the site controllers view their local network and aggregate flows to the higher level centralized controller which performs TE. Though the paper talks about a very particular implementation, it demonstrates in practice the benefits of the generic principle of our design.

@@

\section{Acknowledgements}
\label{sec:ack}

\bibliographystyle{abbrv}
\bibliography{main}

>>>>

\subsection{One switch abstraction}
\label{subsec:aggr}
A controller aggregates its underlying network and behaves as one logical switch, with huge number of ports. This logical switch supports control plane protocols such as OpenFlow\cite{openflow,of-spec}, and can be controlled by an external entity (referred to as parent controller) to install forwarding rules. One switch abstraction has two parts, (i) compressing the underlying network state to represent a single switch, and (ii) translating instructions (forwarding rules) from the abstract switch to the underlying network. We present these separately below.
\subsubsection{Compression}
The primary switch abstraction comes by compressing the network into one node. Given an underlying network for a controller, the abstract switch exposed by it contains all the ports which are connected to switches outside the network. Such ports are referred to as \emph{external} ports. Conversely, ports which are connected to ports in the same network are not exposed (hidden away from abstraction) and are referred to as \emph{internal} ports. \emph{External} ports form the \emph{ingress} and \emph{egress} ports for the abstract switch. The switch abstraction maintains a list of these abstract ports, which can be (uniquely) mapped to the ports of the underlying network. This basic abstraction is sufficient for a simple control application, but we require more features for a rich and complete experience. A few such abstractions are presented below.
\begin{enumerate}
    \item \emph{Host compression}:
    Hosts in the underlying network are considered as external entities, which are exposed as hosts connected to the abstract switch.Ports which connect to hosts are considered to be\emph{external} ports. However, if the number of hosts in the underlying network is too large, exposing a list of all hosts could result in inefficient compression. Host compression can be performed in two ways, depending on the host IP distribution and the network being considered.
    \begin{itemize}
        \item \emph{IP compression} -
        Similar to the forwarding rules in an internet router, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller. This works well only if the assignment of switches to controllers is uniform and in accordance with prefixes.
        \item \emph{Abstract host list} -
        A controller can expose an abstract host or list of hosts to its parent controller. The parent controller places rules for the abstract hosts in the peer controllers. This approach would be similar to LAN, and requires address translation while entering or leaving the underlying network. This is more flexible, but complicated, and can be used only when IP prefix compression fails.
    \end{itemize}
    \item \emph{Abstract flow table}:
        The basic flow table in a controller is comprise of the rules placed by the parent controller. However, for a full OpenFlow functionality, we need to support counters and queries, as in OpenFlow switches. These responses, again, must be an aggregation of the network. To this end, we must be able to efficiently represent the flows from the underlying network and obtain statistics which can be exposed to the parent. We propose a data structure to maintain, aggregate and update view of the flows in section~\ref{subsubsec:aggr-data}.
\end{enumerate}

@@

\cite{one-big-switch} proposes algorithms to combine and translate end-to-end policies and routing policies to optimally place rules in switches. These algorithms can be used by control applications to optimally place rules, combining the abstract flow table and its local routing policies.

@@

\subsection{Hierarchical arrangement}
We propose a new control plane architecture in which controllers are arranged in the form of a tree over the physical network as shown in figure ~\ref{fig:hierarchy}. Switches form the leaf nodes in the tree, while controllers form the internal nodes. An internal node acts as a switch for its parent and as a controller for its children. A controller's view does not span the whole network but only the network of its children. A node translates instructions (forwarding rules) from its parent controller to the forwarding rules for its underlying network. Similarly, network updates from the underlying network are translated into network updates arising from one switch abstraction of the given controller and are passed on to parent controller.

The abstraction described above represents a controller at any level of the hierarchy. The only difference between a root controller and an internal controller is the absence of incoming OpenFLow messages in the former, while the latter receives them from its parent. This simplifies the control application which can be agnostic to the hierarchical level it belongs to, as long as it can support incoming OpenFlow messages.

@@

Hyperflow\cite{hyperflow} has a logically centralized but physically distributed controller architecture. It localizes decision making in individual controllers by passively synchronizing the global network view. The control application is agnostic to the underlying architecture and takes a global decision, though the effect is local. The events are propagated using a publish/subscribe messaging paradigm.

Onix\cite{onix} presents a platform on top of which a network control plane can be implemented as a distributed system. Control planes written within Onix operate on a global view of the network and use basic state distribution primitives provided by the platform. Thus Onix provides a general distributed state management API for control plane implementations, but leaves the decision making and trade-offs among consistency, durability and scalability to the control applications. This makes the design, deployment, modification and verification of control applications difficult and cost prohibitive.

Kandoo\cite{kandoo} takes a slightly similar approach to our own and presents a two layer hierarchy of controllers: (i) the bottom layer is a group of controllers with no interconnection and no knowledge of the nework-wide state, and (ii) the top layer is a logically centralized controller that maintains network-wide state. Bottom layer runs a local application for its underlying network, while the top layer subscribes to changes from the bottom layer. Top layer view is subscription dependent and the architecture cannot be used in a generic fasion.

On a similar but different note, B4\cite{b4} uses a two level hierarchy, where the site controllers view their local network and aggregate flows to the higher level centralized controller which performs TE. Though the paper talks about a very particular implementation, it demonstrates in practice the benefits of the generic principle of our design.

@@

\section{Acknowledgements}
\label{sec:ack}
\nocite{*}

\bibliographystyle{abbrv}
\bibliography{paper}
